---
title: "art-sauce Finance Report"
author: "David Barron"
date: "`r format(Sys.Date(), '%d %B %Y' )`"
output:
  pdf_document: default
  html_document:
    theme:
      bg: '#FFF8DC'
      fg: '#000080'
      primary: '#00FF00'
      base_font: Georgia
      heading_font: Arial
---

```{r setup, include=TRUE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error=TRUE, cache=FALSE)
library(knitr)
library(kableExtra)
options(knitr.kable.NA = '')

# Format money output
inline_hook <- function(x){
  format(x, nsmall = 2, scientific = FALSE, big.mark = ",")
}

knit_hooks$set(inline = inline_hook)

```

```{r intro}
library(tidyverse)
library(googlesheets4)
library(readxl)

```

```{r readdata}
 as_dta <- read_sheet("https://docs.google.com/spreadsheets/d/1SyszVF2SYg594Z6yqFBbyGqJxVR2C6IjfhJF35Obvks/edit?usp=sharing",
   sheet = "Accoounts 2025 January - December")

as_dta <- as_dta %>%
  select(Date, Description, Expenditure = `Debit Amount`, Income = `Credit Amount`, Balance, Category)

nrows <- dim(as_dta)[1]
```

```{r prevdata}
 prev_as_dta <- read_sheet("https://docs.google.com/spreadsheets/d/1eKrWtCjT86qElyHUVfDRuESNdZzGMvXR25UdAMDiTcY/edit?usp=sharing",
   sheet = "Accoounts 2024 January - December")

prev_as_dta <- prev_as_dta %>%
  select(Date, Description, Expenditure = `Debit Amount`, Income = `Credit Amount`, Balance, Category)

prev_nrows <- dim(prev_as_dta)[1]

```

```{r transformdata}

sum_dta <- function(dta){
  dta %>%
          mutate(Date = lubridate::ymd(Date),
                 Month = lubridate::month(Date)) %>%
        group_by(Month) %>%
        summarise(
          Expend = sum(Expenditure, na.rm = TRUE),
          Inc = sum(Income, na.rm = TRUE)
        )  
}
 
as_sum <- sum_dta(as_dta)

calc.cumsum <- function(dta){
  dta %>%
    mutate(
          Income = cumsum(Inc),
          Expenditure = cumsum(Expend)
        ) %>%
        select(Month, Income, Expenditure)
}

as_cumsum <- calc.cumsum(as_sum)

num.months <- dim(as_cumsum)[1]


calc.long <- function(dta) {
  dta %>% pivot_longer(
              cols = Income:Expenditure,
              names_to = "Type",
              values_to = "Amount"
  ) 
}

as_long <- calc.long(as_cumsum) %>%
  mutate(Year = 2025)

calc.sum <- function(dta){
  dta %>%
  group_by(Category) %>%
  summarise(Income = sum(Income, na.rm = TRUE),
            Expenditure = sum(Expenditure, na.rm = TRUE))
}

sum_tab <- calc.sum(as_dta)

total.income <- as.numeric(as_cumsum[num.months, 2])

total.expenditure <- as.numeric(as_cumsum[num.months, 3]) 


```

```{r prevtransform}

prev_as_sum <- sum_dta(prev_as_dta)
prev_as_cumsum <- calc.cumsum(prev_as_sum)

prev_num.months <- dim(prev_as_cumsum)[1]

prev_as_long <- calc.long(prev_as_cumsum) %>% 
  mutate(Year = 2024)

prev_sum_tab <- calc.sum(prev_as_dta)



```

```{r plot}

num.months <- 12 # comment out to only show YTD

labs <- month.abb[1:num.months]

as_full <- rbind(as_long, prev_as_long)

as_full  %>% mutate(Year = factor(Year)) %>%
 # filter(Month <= num.months) %>%
  ggplot(aes(x = Month, y = Amount, colour = Type, linetype = Year)) + 
          geom_point() +
          geom_line() +
          scale_x_continuous(
            breaks = 1:num.months,
            minor_breaks = NULL,
            label =  labs
          ) + 
  scale_y_continuous(label = scales::label_dollar(prefix="£")) +
  scale_linetype_manual(values = c("twodash", "solid")) +
  scale_colour_manual(values = c("red", "green3")) +
  theme_light() +
    theme(panel.background = element_rect(fill = "#FFF8DC"),
        plot.background = element_rect(fill = "#FFF8DC"),
        legend.background = element_rect(fill = "#FFF8DC"))

```

## Income and Expenditure

-   Total income since 1 January 2025 is £`r total.income`
-   Total expenditure since 1 January 2025 is £`r total.expenditure`
-   **The current year's `r ifelse(total.income - total.expenditure > 0, "surplus", "deficit")` is £`r total.income - total.expenditure`**
-   Current account balance £`r as.numeric(as_dta[nrows,5])`

```{r summary}

nrow_sum <- dim(sum_tab)[1]
sum_tab %>% kbl() %>%
  kable_classic(full_width = FALSE, html_font = "Georgia") %>%
    row_spec(0:nrow_sum, color = "#000080") %>%
    row_spec(0, bold = TRUE, color = "white", background = "#000080") #%>%
  #  kable_styling(latex_options = c("hold_position"), position = "center")

```


### Notes

- We have yet to be invoiced by Magdalen Road Studios for the use we have made of MAS
